---
title: "Covtype Exercise"
output:   
  github_document:
    toc: true
    toc_depth: 3
---

```{r}
covtype <- read.csv(file="Data/covtype.data", header=FALSE, sep=",", stringsAsFactors=FALSE)
head(covtype,n = 2)
```


```{r}

# "Fixed" attributes' names
names <- c("Elevation","Aspect","Slope","HorDistToHydro","VertDistToHydro",
           "HorDistRoad","Hillshade09","Hillshade12","Hillshade15",
           "HorDistFire")
# Four binary attributes for wilderness areas:
names <- c(names,"WA_RWA","WA_NWA","WA_CPWA","WA_CLPWA")
# 40 (!) binary attributes for soil types:
names <- c(names,sprintf("ST%02d",1:40))
# The cover type
names <- c(names,"Class")
# Assign these names to the attributes
names(covtype) <- names
# Let's also assign labels to the coverage types.
covtype$Class <- as.factor(covtype$Class)
levels(covtype$Class) <- c("Spruce/Fir", "Lodgepole Pine",
                           "Ponderosa Pine","Cottonwood/Willow","Aspen",
                           "Douglas-fir","Krummholz")
# How does it looks like?
str(covtype)

```



```{r}
# Which are the columns we want to consider?
searchOn <- c("WA_RWA","WA_NWA","WA_CPWA","WA_CLPWA")
# Get the index of each WA_
indexOfWA <- apply(covtype[,searchOn], 1, function(x) which(x == 1))
# Convert it to a factor with the column names we used.
factorOfWA <- factor(indexOfWA,labels = searchOn)
# Add it to the data frame
covtype$WildArea <- factorOfWA
# Drop the binary variables we don't need anymore
covtype[searchOn] <- list(NULL)
```


```{r}
# Which are the columns we want to consider?
searchOn <- sprintf("ST%02d",1:40)

# Get the index of 1 in ST01..ST40
indexOfST <- apply(covtype[,searchOn], 1, function(x) which(x == 1))
# Convert it to a factor with the column names we used.
factorOfST <- factor(indexOfST,labels = searchOn)
# Add it to the data frame
covtype$SoilType <- factorOfST
# Drop the binary variables we don't need anymore
covtype[searchOn] <- list(NULL)
str(covtype)
```


As descrições de tipo de solo podem ser encontradas  [aqui](https://archive.ics.uci.edu/ml/machine-learning-databases/covtype/covtype.info). A tabela com esses dados foi extraída e convertida em um csv. A primeira coluna indica o SoilType e a segunda coluna o *USFS ELU Code*. No *USFS ELU Code*, o primeiro digito representa a zona climática e o segundo a zona geológica. Esses dados também estão aqui [aqui](https://archive.ics.uci.edu/ml/machine-learning-databases/covtype/covtype.info) e foram extraídos para dois arquivos climatic_zone.csv e geologic_zone.csv, respectivamente.

```{r}
soilTypeDesc <- read.csv(file="Data/Soil_Types_Description.csv", header=FALSE, sep=",", stringsAsFactors=FALSE)
names(soilTypeDesc) <- c("SoilType","USFS_ELU_CODE")
head(soilTypeDesc, n=5)
climaticZones  <- read.csv(file="Data/climatic_zones.csv", header=FALSE, sep=",", stringsAsFactors=TRUE)
names(climaticZones) <- c("Digit","ClimaticZone")
head(climaticZones, n=5)
geologicZones  <- read.csv(file="Data/geologic_zones.csv", header=FALSE, sep=",", stringsAsFactors=TRUE)
names(geologicZones) <- c("Digit","GeologicZone")
head(geologicZones, n=5)

soilTypeDesc$ClimateZone <- climaticZones$ClimaticZone[as.integer(soilTypeDesc$USFS_ELU_CODE/1000)]
soilTypeDesc$GeologicZone <- geologicZones$GeologicZone[as.integer((soilTypeDesc$USFS_ELU_CODE - as.integer(soilTypeDesc$USFS_ELU_CODE/1000)*1000)/100)
]

head(soilTypeDesc, n=5)

```


```{r}

covtype$ClimateZone <- soilTypeDesc$ClimateZone[as.integer(covtype$SoilType)]

covtype$GeologicZone <- soilTypeDesc$GeologicZone[as.integer(covtype$SoilType)]

head(covtype[c(13,14,15)], n=5)
```


```{r}
str(covtype)
```

Olhando os dados:

```{r,  results='asis'}
library(xtable)
tabCZ <- table(covtype$ClimateZone,covtype$WildArea)
print(xtable(tabCZ), type="html")
```